Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2014-12-22 doc/GMP2015/s04_recombination.tex
Final changes to main proof and small corrections throughout.

\section{Using the canonical, optimal DR-plan for Realizing (Solving) Qusecs}
\label{sec:recomb}

In this section, we consider the \dfn{optimal recombination} problem
of combining specific solutions of subsystems in a DR-plan into a
solution of their parent system $I$ (without loss of generality, at
the top level of the DR-plan). In the case of isostatic qusecs, the
parent system $I$ is isostatic (the root of the DR-plan), and we seek
\dfn{solution(s) (among a finite large number of solutions) with a
specific orientation or chirality}. In the case of under-constrained
qusecs the subsystems are the multiple roots of the DR-plan, the
parent system $I$ is under-constrained, and we typically seek an
efficient algorithmic description of \dfn{connected component(s) of
solutions with a specific orientation or chirality}.

@@

In the following, we formulate the problem of \dfn{optimal
modification}, of the indecomposable algebraic system $I$ into a
decomposable system with a small DR-plan (low algebraic complexity).
Leveraging recent results on \dfn{Cayley configuration spaces}, our
approach to the optimal modification problem achieves the following:
\dfn{(a) Small DR-plan.} We obtain a  parameterized family of systems
$I_{\lambda_F}$ ---  one for each value $\lambda_F$ for the parameters
$F$,  all of which have small DR-plans. Thus, given a value $v$ for
$\lambda_F$, the system $I_v$ can potentially be solved or realized
easily once the orientation type of the solution is known  (when the
DR-plan size is small enough). \dfn{(b) Solution preservation.}
Moreover, the union of solution spaces of the systems in the family
$I_{\lambda_F}$ is guaranteed to contain all of $I$'s solutions. \dfn{(c) Efficient search.} Finally, this union can be projected into the
so-called \dfn{Cayley} or distance parameter space $\lambda_F$ that is
convex or otherwise easy to traverse in order to search for $I$'s
solution (or connected component) of the desired orientation type. For
the case when the modification is bounded, this approach provides an
efficient algorithm for recombination. We first define the decision
version of the problem of optimal modification for decomposition. The
standard optimization versions are straightforward.

@@

It is immediately clear that indecomposable graphs $G$ that belong in
OMD$_k$ for small $k$ and $s$  lend themselves to modification  into
decomposable graphs satisfying Criteria (a) and (b) above. However, it
is not clear how Criterion (c) is met by OMD graphs.
%
\subsection{Using Convex Cayley Configuration Spaces}
\label{sec:2-tree-reduction}
%
Next we provide the necessary background to describe a specific
approach for achieving the requirements (a)--(c) mentioned above, by
restricting the class of reduced graphs $G' = G\setminus E_1$ and
their isostatic completions $H$ in the above definition of the OMD
problem, and using a key theorem of Convex Cayley configuration spaces
\cite{sitharam2010convex}. This theorem characterizes the class of graphs $H$
and non-edges $F$ (Cayley parameters), such that the set of vectors
$\lambda_F$ of  attainable lengths of the non-edges $F$ is always
convex for any given lengths $\delta$ for the edges of $H$ (i.e, over
all the realizations of the bar-joint constraint system or linkage
$(H,\delta)$ in 2 dimensions). This set is called the (2-dimensional)
\dfn{Cayley configuration space} of the linkage $(H,\delta)$ on the
Cayley parameters $F$, denoted $\Phi_F(H,\delta)$ and can be viewed as
a ``projection'' of the Cartesian realization space of $(H,\delta)$ on
the Cayley parameters $F$. Such graphs $H$ are said to have
\dfn{convexifiable Cayley configuration spaces with parameters $F$}
(short: \dfn{convexifiable}).

To state the theorem, we first have to
define the notion of \dfn{2-sums} and \dfn{2-trees}.
Let $H_1$ and $H_2$ be two graphs on disjoint sets of vertices $V_1$ and $V_2$, with edge sets $E_1$ and $E_2$ containing edges $(u,v)$ and $(w,x)$ respectively.
% Let $H_1=(V_1,E_1)$ and $H_2=(V_2,E_2)$ be two graphs on disjoint sets of vertices that contain the edges $(u,v)$ and $(w,x)$ respectively.
A \dfn{2-sum} of $H_1$ and $H_2$ is a graph $H$
obtained by taking the union of $H_1$ and $H_2$ and identifying $u=w$
and $v=w$. In this case, $H_1$ and $H_2$ are called \dfn{2-sum
components} of $H$. A \dfn{minimal 2-sum component} of $H$ is  one
that cannot be further split into 2-sum components. A \dfn{2-tree} is
recursively obtained by taking a 2-sum of 2-trees, with the base
case being a triangle. A \dfn{partial 2-tree} is a 2-tree
minus some edges. Partial 2-trees characterized as having $K_4$ as a
forbidden minor are also called series parallel graphs.

@@

% \begin{proof}
%     The Cartesian realization space of $(H,\left<\delta_{E'},
%     \lambda_F\right>)$ is computed easily with a DR-plan of size 2,
%     and is the union of $2^t$ solutions (modulo orientation preserving
%     isometries) each with a distinct orientation type, where $t$ is
%     the number of triangles in the 2-tree $H$; here $\delta_{S}$ is
%     the restriction of the length vector $\delta$ to the edges in $S$.
%     A desired solution $p$ (or connected component of a solution
%     space) of $(G,\delta)$ of an orientation type $\sigma_p$ can be
%     found by a subdivided binary search   of the cartesian realization
%     space of $(H, \left<\delta_{E'}, \lambda_F\right>)$ of orientation
%     type $\sigma_p$ - as $\lambda_F$ ranges over the discretized
%     convex polytope $\Phi_F(G',\delta_E')$ with bounding hyperplanes
%     described in Theorem \ref{theorem:convexcayley}. A solution $p$
%     is found  when the lengths for nonedges in $D$ match $\delta_D$.
% \end{proof}

\begin{figure*}\centering
\begin{subfigure}{.2\linewidth}\centering
    \input{img/tikz/omd_k33_example_new-0.tikz}
    \caption{}\label{fig:omd_k33_example:a}
\end{subfigure}\hfill
\begin{subfigure}{.15\linewidth}\centering
    \input{img/tikz/omd_k33_example_new-1.tikz}
    \caption{}\label{fig:omd_k33_example:b}
\end{subfigure}
\begin{subfigure}{.15\linewidth}\centering
    \input{img/tikz/omd_k33_example_new-2.tikz}
    \caption{}\label{fig:omd_k33_example:c}
\end{subfigure}\hfill
\begin{subfigure}{.15\linewidth}\centering
    \input{img/tikz/omd_k33_example-1.tikz}
    \caption{}\label{fig:omd_k33_example:d}
\end{subfigure}
\begin{subfigure}{.15\linewidth}\centering
    \input{img/tikz/omd_k33_example-2.tikz}
    \caption{}\label{fig:omd_k33_example:e}
\end{subfigure}

@@

\noindent
\note A huge advantage of the convex Cayley method is that it is
completely unaffected when $\delta$ are intervals rather than exact
values~\cite{sitharam2010convex}.

\myexample
A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial
isostatic graphs, i.e, its DR-plan has a root and 9 children
corresponding to the 9 edges. Solving or recombining the system
$(G,\delta)$ corresponding to the root of this DR-plan involves
solving a quadratic system with 8 equations and variables. Instead of
simultaneously solving this system, we could instead use the fact that
$G=K_{3,3}$ is in OMD$_1$: remove the edge $e$ in Figure
\ref{fig:omd_k33_example} to give a partial 2-tree $G'$. Now add the
non-edge $f$ to give a 2-tree $H$ with a DR-plan of size 2. The Cayley
configuration space $\Phi_f(G', \delta_{E\setminus e})$ is a single
interval of attainable length values $\lambda_F$ for the edge $f$.
When $\delta$ is generic, i.e, does not admit collinearities or
coincidences in the realizations of $(G,\delta)$, the realization
space of $(H, \left<\delta_{E\setminus e}, \lambda_f\right>)$ has 16
solutions $q^p_{\lambda_f}$ (modulo orientation preserving
isometries), with distinct orientation types $\sigma_p$ (two
orientation choices for each of the 4 triangles) that can be obtained
by solving a sequence of 4 single quadratics in 1 variable (DR-plan of
size 2). By subdivided binary search in the interval $\lambda_f \in
\Phi_f(G', \delta_{E\setminus e})$, the desired solution $p$  of
$(G,\delta)$ is found when the length of the non-edge $e$ in  the
realization $q^p_{\lambda_f}$ is $\delta_e$.
%
\subsection{Optimized Modification by Enlarging the class of Reduced
Graphs}
\label{sec:tdecomp}
%
It is possible in principle to decrease $k$ for a OMD$_k$ graph (i.e,
the number of edges to be removed to ensure an isostatic completion
that is decomposable with a small DR-plan) by considering reduced
graphs $G'$ (and modified graphs $H$) that come from a larger class
than partial 2-trees but nevertheless have convex Cayley configuration
spaces at least when the realization space is restricted to a
sufficiently comprehensive orientation type. In particular, the so-
called \dfn{tree-decomposable graphs of low Cayley complexity}
\cite{sitharam2011cayleyI,sitharam2011cayleyII} include the partial 2-trees and many others that are not
partial 2-trees. These too result in DR-plans of size 2 or 3, putting
$G$ in the class OMD$_k$ and thus meeting Criteria (a) and (b). The
Criterion (c) is met --- for example when $k=1$ --- because 1-dof Cayley
configuration spaces of linkages based on such graphs $G'$ are known
to be single intervals when a comprehensive orientation type
$\sigma_p$ of the sought solution $p$ is given. In addition, the
bounds of these intervals are of low algebraic complexity. More
precisely, the bounds  can themselves be computed using a DR-plan of
size 2 or 3, i.e, the computation of these bounds is tree-
decomposable. Alternatively, the bounds are in a simple quadratic or
radically solvable extension field over the rationals, or can be
computed by solving a triangularized system of quadratics.

>>>>

\section{Using the canonical, optimal DR-plan for Realizing (Solving) Qusecs}
\label{sec:recomb}
%
In this section, we consider the \dfn{optimal recombination} problem
of combining specific solutions of subsystems in a DR-plan into a
solution of their parent system $I$ (without loss of generality, at
the top level of the DR-plan). In the case of isostatic qusecs, the
parent system $I$ is isostatic (the root of the DR-plan), and we seek
\vemph{solution(s) (among a finite large number of solutions) with a
specific orientation or chirality}. In the case of underconstrained
qusecs the subsystems are the multiple roots of the DR-plan, the
parent system $I$ is underconstrained, and we typically seek an
efficient algorithmic description of \vemph{connected component(s) of
solutions with a specific orientation or chirality}.

@@

In the following, we formulate the problem of \dfn{optimal
modification}, of the indecomposable algebraic system $I$ into a
decomposable system with a small DR-plan (low algebraic complexity).
Leveraging recent results on \dfn{Cayley configuration spaces}, our
approach to the optimal modification problem achieves the following:

\medskip\noindent
\header{(a) Small DR-plan.} We obtain a  parameterized family of systems
$I_{\lambda_F}$ --- one for each value $\lambda_F$ for the parameters
$F$,  all of which have small DR-plans. Thus, given a value $v$ for
$\lambda_F$, the system $I_v$ can potentially be solved or realized
easily once the orientation type of the solution is known  (when the
DR-plan size is small enough).

\medskip\noindent
\header{(b) Solution preservation.}
Moreover, the union of solution spaces of the systems in the family
$I_{\lambda_F}$ is guaranteed to contain all of $I$'s solutions.

\medskip\noindent
\header{(c) Efficient search.} Finally, this union can be projected into the
so-called \dfn{Cayley} or distance parameter space $\lambda_F$ that is
convex or otherwise easy to traverse in order to search for $I$'s
solution (or connected component) of the desired orientation type. For
the case when the modification is bounded, this approach provides an
efficient algorithm for recombination. We first define the decision
version of the problem of optimal modification for decomposition. The
standard optimization versions are straightforward.

@@

It is immediately clear that indecomposable graphs $G$ that belong in
OMD$_k$ for small $k$ and $s$  lend themselves to modification  into
decomposable graphs satisfying Criteria (a) and (b) above. However, it
is not clear how Criterion (c) is met by OMD graphs.
%
\subsection{Using Convex Cayley Configuration Spaces}
\label{sec:2-tree-reduction}
%
Next we provide the necessary background to describe a specific
approach for achieving the requirements (a)--(c) mentioned above, by
restricting the class of reduced graphs $G' = G\setminus E_1$ and
their isostatic completions $H$ in the above definition of the OMD
problem, and using a key theorem of Convex Cayley configuration spaces
\cite{sitharam2010convex}. This theorem characterizes the class of graphs $H$ and non-edges $F$ (Cayley parameters), such that the set of vectors
$\lambda_F$ of  attainable lengths of the non-edges $F$ is always
convex for any given lengths $\delta$ for the edges of $H$ (i.e, over
all the realizations of the bar-joint constraint system or linkage
$(H,\delta)$ in 2 dimensions). This set is called the (2-dimensional)
\dfn{Cayley configuration space} of the linkage $(H,\delta)$ on the
Cayley parameters $F$, denoted $\Phi_F(H,\delta)$ and can be viewed as
a ``projection'' of the cartesian realization space of $(H,\delta)$ on
the Cayley parameters $F$. Such graphs $H$ are said to have
\dfn{convexifiable Cayley configuration spaces with parameters $F$}
(short: \dfn{convexifiable}).

To state the theorem, we first have to define the notion of
\dfn{2-sums} and \dfn{2-trees}. Let $H_1$ and $H_2$ be two graphs on
disjoint sets of vertices $V_1$ and $V_2$, with edge sets $E_1$ and
$E_2$ containing edges $(u,v)$ and $(w,x)$ respectively.% Let $H_1=(V_1,E_1)$ and $H_2=(V_2,E_2)$ be two graphs on disjoint sets of vertices that contain the edges $(u,v)$ and $(w,x)$ respectively.
A \dfn{2-sum} of $H_1$ and $H_2$ is a graph $H$ obtained by taking the
union of $H_1$ and $H_2$ and identifying $u=w$ and $v=w$. In this
case, $H_1$ and $H_2$ are called \dfn{2-sum components} of $H$. A
\dfn{minimal 2-sum component} of $H$ is  one that cannot be further
split into 2-sum components. A \dfn{2-tree} is recursively obtained by
taking a 2-sum of 2-trees, with the base case of a 2-tree being a
triangle. A \dfn{partial 2-tree} is a 2-tree minus some edges. Partial
2-trees have an alternate characterization as the graphs that avoid
$K_4$ minors, and are also called series parallel graphs.

@@

% \begin{proof}
%     The Cartesian realization space of $(H,\left<\delta_{E'},
%     \lambda_F\right>)$ is computed easily with a DR-plan of size 2,
%     and is the union of $2^t$ solutions (modulo orientation preserving
%     isometries) each with a distinct orientation type, where $t$ is
%     the number of triangles in the 2-tree $H$; here $\delta_{S}$ is
%     the restriction of the length vector $\delta$ to the edges in $S$.
%     A desired solution $p$ (or connected component of a solution
%     space) of $(G,\delta)$ of an orientation type $\sigma_p$ can be
%     found by a subdivided binary search   of the cartesian realization
%     space of $(H, \left<\delta_{E'}, \lambda_F\right>)$ of orientation
%     type $\sigma_p$ - as $\lambda_F$ ranges over the discretized
%     convex polytope $\Phi_F(G',\delta_E')$ with bounding hyperplanes
%     described in Theorem \ref{theorem:convexcayley}. A solution $p$
%     is found  when the lengths for nonedges in $D$ match $\delta_D$.
% \end{proof}
%
\begin{figure*}\centering
\begin{subfigure}{.2\linewidth}\centering
    \input{img/tikz/omd_k33_example_new-0.tikz}
    \caption{}\label{fig:omd_k33_example:a}
\end{subfigure}\hfill
\begin{subfigure}{.15\linewidth}\centering
    \input{img/tikz/omd_k33_example_new-1.tikz}
    \caption{}\label{fig:omd_k33_example:b}
\end{subfigure}
\begin{subfigure}{.15\linewidth}\centering
    \input{img/tikz/omd_k33_example_new-2.tikz}
    \caption{}\label{fig:omd_k33_example:c}
\end{subfigure}\hfill
\begin{subfigure}{.15\linewidth}\centering
    \input{img/tikz/omd_k33_example-1.tikz}
    \caption{}\label{fig:omd_k33_example:d}
\end{subfigure}
\begin{subfigure}{.15\linewidth}\centering
    \input{img/tikz/omd_k33_example-2.tikz}
    \caption{}\label{fig:omd_k33_example:e}
\end{subfigure}

@@

\noindent
\note A major advantage of the convex Cayley method is that it is
completely unaffected when $\delta$ are intervals rather than exact
values~~\cite{sitharam2010convex}.
%cite my paper with Heping!!!

\myexample
[Using Cayley configuration space].
A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial
isostatic graphs, i.e, its DR-plan has a root and 9 children
corresponding to the 9 edges. Solving or recombining the system
$(G,\delta)$ corresponding to the root of this DR-plan involves
solving a quadratic system with 8 equations and variables. Instead of
simultaneously solving this system, we could instead use the fact that
$G=K_{3,3}$ is in OMD$_2$: remove the edges $e_1,e_2$ in Figure
\ref{fig:omd_k33_example} to give a partial 2-tree $G'$. Now add the
nonedges $f_1,f_2$ to give a 2-tree $H$ with a DR-plan of size 2. The Cayley
configuration space $\Phi_f(G', \delta_{E\setminus e})$ is a single
interval of attainable length values $\lambda_F$ for the edge $f$.
When $\delta$ is generic, i.e, does not admit collinearities or
coincidences in the realizations of $(G,\delta)$, the realization
space of $(H, \left<\delta_{E\setminus e}, \lambda_f\right>)$ has 16
solutions $q^p_{\lambda_f}$ (modulo orientation preserving
isometries), with distinct orientation types $\sigma_p$ (two
orientation choices for each of the 4 triangles) that can be obtained
by solving a sequence of 4 single quadratics in 1 variable (DR-plan of
size 2). By subdivided binary search in the interval $\lambda_f \in
\Phi_f(G', \delta_{E\setminus e})$, the desired solution $p$  of
$(G,\delta)$ is found when the length of the nonedge $e$ in  the
realization $q^p_{\lambda_f}$ is $\delta_e$.

In fact, we can show that $G=K_{3,3}$ is in OMD$_1$ by removing a single
edge to reduce (as shown in Figure \ref{fig:omd_k33_example}) to a
tree-decomposable graph of low Cayley complexity (which includes the class
of partial 2-trees).
In the next Section \ref{sec:tdecomp}, we discuss this issue
of why the largest class of reduced graphs is desirable.
%
\subsection{Optimized Modification by Enlarging the class of Reduced
Graphs}
\label{sec:tdecomp}
%
It is possible in principle to decrease $k$ for a OMD$_k$ graph (i.e,
the number of edges to be removed to ensure an isostatic completion
that is decomposable with a small DR-plan) by considering reduced
graphs $G'$ (and modified graphs $H$) that come from a larger class
than partial 2-trees but nevertheless have convex Cayley configuration
spaces at least when the realization space is restricted to a
sufficiently comprehensive orientation type. In particular, the so-
called \dfn{tree-decomposable graphs of low Cayley complexity}
\cite{sitharam2011cayleyI,sitharam2011cayleyII} include the partial 2-trees and many others that are not
partial 2-trees. See an illustration in Figure \ref{fig:omd_k33_example}.
These too result in DR-plans of size 2 or 3, putting
$G$ in the class OMD$_k$ and thus meeting Criteria (a) and (b). The
Criterion (c) is met --- for example when $k=1$ --- because 1-dof Cayley
configuration spaces of linkages based on such graphs $G'$ are known
to be single intervals when a comprehensive orientation type
$\sigma_p$ of the sought solution $p$ is given. In addition, the
bounds of these intervals are of low algebraic complexity. More
precisely, the bounds  can themselves be computed using a DR-plan of
size 2 or 3, i.e, the computation of these bounds is tree-
decomposable. Alternatively, the bounds are in a simple quadratic or
radically solvable extension field over the rationals, or can be
computed by solving a triangularized system of quadratics.

### 2014-12-22 doc/GMP2015/s08_pinnedline.tex
Meera's changes to 1,2,3,8. Minor changes to refs and cites.
Final changes to main proof and small corrections throughout.

A \dfn{pinned line-incidence system} $(G,\delta)$ is a graph $G=(V,E)$
together with parameters $\delta$ specifying $|E|$ \textbf{pins} with
fixed positions in $\mathbb{R}^2$, such that each edge is constrained
to lie on a line passing through the corresponding pin, i.e.\ $\delta:
E \rightarrow \mathbb{R}^2$.
%
% $G=(V,E, w_V, w_E)$
%%is a \textbf{2-dimensional pinned line-incidence graph},
% has weight functions $w_V(v) = 2, w_E(e) = 1$ for all $v \in V, e \in E$.
%Additionally, there is  a  set $X$ of $|E|$ \textbf{pins} with fixed positions in $\mathbb{R}^2$,
%such that  %every edge $e \in E$ corresponds to a pin $x \in X$ where
%for each edge $e \in E$, there is a uniquely corresponding pin $x \in X$,
%such that $e$ is constrained to lie on a line passing through $x$.
%%There exists a set $X$ of $|E|$ pins fixed in $\mathbb{R}^2$, such that each edge $e_i$
%
A pinned line-incidence graph $G$ is rigid if $|E| = 2|V|$ and $|E'| \le 2|V'|$ for every induced subgraph $(V',E')$ \cite{sitharam2014incidence}. Note that no trivial motion exists since the pins have fixed positions on the plane.
Euclidean transformations are not factored out.
In particular, both a single vertex and a single edge are under-constrained graphs.

@@

In this section, we will adapt the results in Section~\ref{sec:DRP}
to give the canonical DR-plan for pinned line-incidence graphs.
First, we note that %unlike bar-joint graphs,
\vemph{an isostatic pinned line-incidence graph can be disconnected},
being the disjoint union of two or more isostatic subgraphs.
This is because the pins have fixed positions on the plane.
%
%the following difference of pinned line-incidence graphs from the bar-joint graphs.
%
%For pinned line-incidence graphs,
%since the pins have fixed positions on the plane,
%{\em a well-constrained vertex-maximal proper subgraph can be disconnected},
%being the disjoint union of two or more well-constrained subgraphs.
%
%\begin{itemize}
%
%\item For pinned line-incidence graphs,
%since the pins have fixed positions on the plane,
%no trivial motion exists, therefore {\em no trivial over-constrained graphs exist} (the smallest over-constrained graph has $6$ vertices).\\
%In particular, note that both a single vertex and a single edge are under-constrained graphs.
%
%\item
%For pinned line-incidence graphs,
%since the pins have fixed positions on the plane,
%{\em a well-constrained vertex-maximal proper subgraph can be disconnected},
%being the disjoint union of two or more well-constrained subgraphs. \\
%%Without loss of generality,
%%we slightly modify the requirement of DR-plan for pinned line-incidence graphs
%%such that only connected subgraphs can be considered as a node of the DR-plan.
%
%\end{itemize}
%
We define a \dfn{trivial} graph to be a single vertex and make the following modification to the definition of 
the canonical DR-plan:

%\begin{itemize}
%\item For any $C$, its $N$ children, $C_1, \ldots, C_N$, are {\em edges} or {\em connected} well-constrained vertex-maximal proper subgraphs of that node $C$.
%\item A node containing a single edge is a leaf.
%%\item
%%\todo{We just define the %trivial graph/
%%DR-plan leaves for 2-dimensional pinned line-incidence graphs to be single vertices...  }
%%\item \todo{A single edge can be a node of the DR-plan}
%%\item only connected subgraphs can be considered as a node of the DR-plan.
%\end{itemize}

\begin{definition}
The \dfn{DR-plan} of a  pinned line-incidence graph $G$ is one in which
(1) each child node of a non-leaf node $C$ is either a \dfn{connected} rigid vertex-induced subgraph of $C$,
or an edge not contained in any  proper rigid subgraph of $C$, and (2) a leaf node is a single edge.
% Definition~\ref{def:drp} with  modified rules number 2 and 4:
%    The \textbf{decomposition-recombination (DR-) plan} of  a 2-dimensional pinned line-incidence graph $G$, $DRP(G)$, is  a tree that has the following properties:
%    \begin{enumerate}
%        %\item The root of the tree `contains' $G$.
%        \item[2] For any non-leaf node $C$, each of its children, $C_1,\ldots,C_N$, is either an edge not contained in any well-constrained proper subgraph of $C$, or a {\em connected} well-constrained vertex-maximal proper subgraph of  $C$.
%        %\item The vertex set of $\bigcup_{i=1}^N{C_i}$ is the vertex set of $C$.
%        \item[4] A node containing a single edge is a leaf.
%    \end{enumerate}

@@

Theorem~\ref{theorem:canonical_is_optimal} and \ref{theorem:main}
hold for  pinned line-incidence graphs with the above modified definition. 
The proof is similar to the original proof (in the Appendix) using the same set of lemmas and  
%
%\begin{corollary}\label{cor:pinned}
%Given a isostatic pinned line-incidence graph, for node $C$ in $OptimalDRP(G)$ and the children of $C$ in $CompleteDRP(C)$ labeled as $C_1,\ldots,C_N$
%\begin{enumerate}
%    \item if $C_i \cap C_j$ is trivial then all $C_1,\ldots,C_N$ are children of $C$ in $OptimalDRP(G)$.
%    \item if $C_i \cap C_j$ is isostatic then any two out of $C_1,\ldots,C_N$ will be the only children of $C$ in $OptimalDRP(G)$.
%\end{enumerate}
%\end{corollary}
%
the following modified version of Observation~\ref{lemma:union_intersection},
% observation, which
%serves a similar function to Observation~\ref{lemma:union_intersection} in Section~\ref{sec:DRP}
%Using similar counting-based argument, %  to Lemma~\ref{t-l1}--\ref{iuc-l1} in Section~\ref{sec:union_intersection},
which  can be proved using a simple counting based argument.

\begin{observation}\label{lem:pinned_union_intersection}
Let $F_i$ and $F_j$ to be subgraphs of the same isostatic graph $F$,
where each of them can be either a single edge or a connected isostatic subgraph.
There are only two possible cases:
(1) At least one of $F_i$, $F_j$ is an edge, if and only if $F_i \cup F_j$ is under-constrained, if and only if $F_i \cap F_j$ is trivial; and (2) both $F_i$ and $F_j$ are isostatic, if and only if $F_i \cup F_j$ is isostatic, if and only if $F_i \cap F_j$ is isostatic.
%The following table summarizes all bijections:
%\setlength{\tabcolsep}{3pt}
%{\small
%\begin{center}
%\begin{tabular}{|p{2.8cm}|c|p{2.3cm}|}
%\hline
%\textbf{Property of $F_i,F_j$} & \textbf{$F_i\cup F_j$} & \centering \textbf{$F_i\cap F_j$}  \tabularnewline \hline
% \centering--- &  \sout{over-constrained}          &  \centering --- \tabularnewline \hline
%at least one of $F_i,F_j$ is an edge & under-constrained       &  \centering a vertex (under-constrained)  \tabularnewline \hline
%both $F_i$ and $F_j$ are well-constrained & well-constrained        &  \centering well-constrained  \tabularnewline \hline
% \centering --- & ---                     &  \centering \sout{under-constrained} \tabularnewline \hline
%\end{tabular}
%\end{center}
%}
%\setlength{\tabcolsep}{6pt}
\end{observation}

>>>>

A \dfn{pinned line-incidence system} $(G,\delta)$ is a graph $G=(V,E)$
together with parameters $\delta$ specifying $|E|$ \textbf{pins} with
fixed positions in $\mathbb{R}^2$, such that each edge is constrained
to lie on a line passing through the corresponding pin, i.e.\ $\delta:
E \rightarrow \mathbb{R}^2$.
%
% $G=(V,E, w_V, w_E)$
%%is a \textbf{2-dimensional pinned line-incidence graph},
% has weight functions $w_V(v) = 2, w_E(e) = 1$ for all $v \in V, e \in E$.
%Additionally, there is  a  set $X$ of $|E|$ \textbf{pins} with fixed positions in $\mathbb{R}^2$,
%such that  %every edge $e \in E$ corresponds to a pin $x \in X$ where
%for each edge $e \in E$, there is a uniquely corresponding pin $x \in X$,
%such that $e$ is constrained to lie on a line passing through $x$.
%%There exists a set $X$ of $|E|$ pins fixed in $\mathbb{R}^2$, such that each edge $e_i$
%
A pinned line-incidence graph $G$ is rigid if $|E| = 2|V|$ and $|E'| \le 2|V'|$ for every induced subgraph $(V',E')$ \cite{sitharam2014incidence}. Note that no trivial motion exists since the pins have fixed positions on the plane.
Euclidean transformations are not factored out.
In particular, both a single vertex and a single edge are underconstrained graphs.

@@

In this section, we will adapt the results in Section~\ref{sec:DRP}
to give the canonical DR-plan for pinned line-incidence graphs.
First, we note that %unlike bar-joint graphs,
\vemph{an isostatic pinned line-incidence graph can be disconnected},
being the disjoint union of two or more isostatic subgraphs.
This is because the pins have fixed positions on the plane.
%
%the following difference of pinned line-incidence graphs from the bar-joint graphs.
%
%For pinned line-incidence graphs,
%since the pins have fixed positions on the plane,
%{\em a wellconstrained vertex-maximal proper subgraph can be disconnected},
%being the disjoint union of two or more wellconstrained subgraphs.
%
%\begin{itemize}
%
%\item For pinned line-incidence graphs,
%since the pins have fixed positions on the plane,
%no trivial motion exists, therefore {\em no trivial overconstrained graphs exist} (the smallest overconstrained graph has $6$ vertices).\\
%In particular, note that both a single vertex and a single edge are underconstrained graphs.
%
%\item
%For pinned line-incidence graphs,
%since the pins have fixed positions on the plane,
%{\em a wellconstrained vertex-maximal proper subgraph can be disconnected},
%being the disjoint union of two or more wellconstrained subgraphs. \\
%%Without loss of generality,
%%we slightly modify the requirement of DR-plan for pinned line-incidence graphs
%%such that only connected subgraphs can be considered as a node of the DR-plan.
%
%\end{itemize}
%
We define a \dfn{trivial} graph to be a single vertex and make the following modification to the definition of
the canonical DR-plan:

%\begin{itemize}
%\item For any $C$, its $N$ children, $C_1, \ldots, C_N$, are {\em edges} or {\em connected} wellconstrained vertex-maximal proper subgraphs of that node $C$.
%\item A node containing a single edge is a leaf.
%%\item
%%\todo{We just define the %trivial graph/
%%DR-plan leaves for 2-dimensional pinned line-incidence graphs to be single vertices...  }
%%\item \todo{A single edge can be a node of the DR-plan}
%%\item only connected subgraphs can be considered as a node of the DR-plan.
%\end{itemize}

\begin{definition}
The \dfn{DR-plan} of a  pinned line-incidence graph $G$ is one in which
(1) each child node of a non-leaf node $C$ is either a \dfn{connected} rigid vertex-induced subgraph of $C$,
or an edge not contained in any  proper rigid subgraph of $C$, and (2) a leaf node is a single edge.
% Definition~\ref{def:drp} with  modified rules number 2 and 4:
%    The \textbf{decomposition-recombination (DR-) plan} of  a 2-dimensional pinned line-incidence graph $G$, $DRP(G)$, is  a tree that has the following properties:
%    \begin{enumerate}
%        %\item The root of the tree `contains' $G$.
%        \item[2] For any non-leaf node $C$, each of its children, $C_1,\ldots,C_N$, is either an edge not contained in any wellconstrained proper subgraph of $C$, or a {\em connected} wellconstrained vertex-maximal proper subgraph of  $C$.
%        %\item The vertex set of $\bigcup_{i=1}^N{C_i}$ is the vertex set of $C$.
%        \item[4] A node containing a single edge is a leaf.
%    \end{enumerate}

@@

Theorem~\ref{theorem:main}
holds for  pinned line-incidence graphs with the above modified definition.
The proof is similar to the original proof (in the Appendix) using the same set of lemmas and
%
%\begin{corollary}\label{cor:pinned}
%Given a isostatic pinned line-incidence graph, for node $C$ in $OptimalDRP(G)$ and the children of $C$ in $CompleteDRP(C)$ labeled as $C_1,\ldots,C_N$
%\begin{enumerate}
%    \item if $C_i \cap C_j$ is trivial then all $C_1,\ldots,C_N$ are children of $C$ in $OptimalDRP(G)$.
%    \item if $C_i \cap C_j$ is isostatic then any two out of $C_1,\ldots,C_N$ will be the only children of $C$ in $OptimalDRP(G)$.
%\end{enumerate}
%\end{corollary}
%
the following modified version of Observation~\ref{lemma:union_intersection},
% observation, which
%serves a similar function to Observation~\ref{lemma:union_intersection} in Section~\ref{sec:DRP}
%Using similar counting-based argument, %  to Lemma~\ref{t-l1}--\ref{iuc-l1} in Section~\ref{sec:union_intersection},
which  can be proved using a simple counting based argument.

\begin{observation}\label{lem:pinned_union_intersection}
Let $F_i$ and $F_j$ to be subgraphs of the same isostatic graph $F$,
where each of them can be either a single edge or a connected isostatic subgraph.
There are only two possible cases:
(1) At least one of $F_i$, $F_j$ is an edge, if and only if $F_i \cup F_j$ is underconstrained, if and only if $F_i \cap F_j$ is trivial; and (2) both $F_i$ and $F_j$ are isostatic, if and only if $F_i \cup F_j$ is isostatic, if and only if $F_i \cap F_j$ is isostatic.
%The following table summarizes all bijections:
%\setlength{\tabcolsep}{3pt}
%{\small
%\begin{center}
%\begin{tabular}{|p{2.8cm}|c|p{2.3cm}|}
%\hline
%\textbf{Property of $F_i,F_j$} & \textbf{$F_i\cup F_j$} & \centering \textbf{$F_i\cap F_j$}  \tabularnewline \hline
% \centering--- &  \sout{overconstrained}          &  \centering --- \tabularnewline \hline
%at least one of $F_i,F_j$ is an edge & underconstrained       &  \centering a vertex (underconstrained)  \tabularnewline \hline
%both $F_i$ and $F_j$ are wellconstrained & wellconstrained        &  \centering wellconstrained  \tabularnewline \hline
% \centering --- & ---                     &  \centering \sout{underconstrained} \tabularnewline \hline
%\end{tabular}
%\end{center}
%}
%\setlength{\tabcolsep}{6pt}
\end{observation}
